Quantitative quality of lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations

### 2017-03-29 MANUSCRIPT/manuscript.tex
Update manuscript tex

@@

@@

\section{comparison of acyl chain order parameters between experiments and simulations}
Order parameters from simulations and experiments for acyl chains of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC)
are shown in Fig. \ref{OrderParametersCHOL}.
 \begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}

  \caption{\label{OrderParametersCHOL}
    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).}

  \todo{Why are the order parameters of CHARMM36 too large compared to simulations even without cholesterol? 
    Discussion in https://github.com/NMRLipids/NmrLipidsCholXray/issues/4}

  \todo{Why there is decrease in order parameters towards beginning of acyl chain in MacRog model without cholesterol?}

  \todo{Do the results suggests that condensation effect is too strong in MacRog and Berger models?
  Discussion in https://github.com/NMRLipids/NmrLipidsCholXray/issues/5}
\end{figure*}

The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
(currently only sn-1).

 \begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/OrderParametersCHANGEScholSN1.eps}

  \caption{\label{OrderParametersCHOLchanges}
    Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.}

\newpage
\section{Comparison of form factors between experiments and simulations}
The form factors calculated from different simulations with different cholesterol content are shown in 
Fig.~\ref{FormFactors}.

 \begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/FormFactors.eps}

  \caption{\label{FormFactors}
    Form factors from simulations and experiments.
  }

>>>>

\section{Results and Discussion}

\subsection{Structural sampling of molecules in bilayers with cholesterol}
Order parameters for C-H bond vectors in lipid bilayer systems, measured
with 13C or 2H NMR techniques, give indirect information about structural
sampling of indidual molecules \cite{ollila16}. The observed increase of
acyl chain order parameters with added cholesterol in simulations and
experiments can be explained by increased trans conformations in acyl chains \cite{??},
which is suggested to play critical role in phase behaviour in PC lipid-cholesterol
mixtures \cite{??}. Consequently, the correct cholesterol ordering effect is
expected to be necessary condition for a model used to understand lipid-cholesterol
phase behaviour.

Acyl chain order parameters for pure 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
and mixture with 50 mol\% of cholesterol from different simulations and experiments
are shown in Fig. \ref{OrderParametersCHOL}. Phase separation is not observed for this system \cite{ferreira13,??},
thus the intermolecular interactions are expected to have maximum effect in equimolar mixture.
Experimental acyl chain order parameters are typically well reproduced in state of the art lipid models
for pure lipid bilayers (for review see \cite{ollila16}). This is also mainly observed in the data
presented here, except that CHARMM36 slightly overestimates the order parameters and MACROG model
underestimates order parameters in the beginning of sn-1 chain. In equimolar mixtures with cholesterol
all models overestimate the order parameters. For Berger/Holtje and MACROG models the overestimation is significant,
while less severe for CHARMM36 and Slipid models.\\
\todo{Is overestimation in CHARMM36 and Slipid significant or not?} \\
\todo{Lipid14 results from recent article should be added in this figure} 
\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOL.eps}
  \caption{\label{OrderParametersCHOL}
    Order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
\end{figure*}

The order parameter changes as a function of cholesterol for each segment are shown in Fig.~\ref{OrderParametersCHOLchanges}
(currently only sn-1). \\
\todo{This figure should be finished and similar for sn-2 should be done as well}
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/OrderParametersCHANGEScholSN1.eps}
  \caption{\label{OrderParametersCHOLchanges}
    Order parameter changes from simulations and experiments for each segment in sn-1 chain of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC) as a function of cholesterol concentration.}

\todo{Also experimental cholesterol order parameters are available. Maybe these should be calculated from simulations as well.}

\subsection{Lipid bilayer dimensions and density profiles as a function of cholesterol}
Cholesterol induced changes in lipid packing, thickness and density profiles are suggested
to be relevant, for example, for membrane protein interactions and cellular membrane
lateral organization \cite{??}. Experimental and simulation studies show that
the cholesterol induces the so called ''condensing effect'', i.e. acyl chain ordering
increases the membrane thickness and
reduces the area per molecule in lipid bilayers \cite{??}. This is demonstrated
from simulations in Fig. \ref{apls}, showing the area per PC headgroups and per
total number of molecules as a function of cholesterol concentration from different
models. All models show reduction of area per total amount of molecules, partly
due to the smaller area covered by the cholesterol than lipids, but also because
lipids ordered by cholesterol require less space. Due to the latter effect,
the area per PC headgroup do not essentially increase up to the addition of
$\sim$15 mol\% of cholesterol. \\
\todo{Membrane thickness to be plotted as well.}
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/apls.eps}
  \caption{\label{apls}
    Area per molecules calculated from different simulation models
    as a function of cholesterol concentration.
  }
\end{figure}

Bilayer dimensions are indirectly experimentally accessible by scattering methods 
and have been widely used to validate lipid bilayer simulations \cite{??}. 
While sophisticated models are applied to extract area per molecule and thicknes for single
component lipid bilayers, the results for multicomponent systems are more difficult to interpret \cite{??}.
In such case the validation can be done by comparing directly measurable form factors between
simulations and experiments \cite{??}. If simulation reproduces the form factor, it can be also
used to interpret the experiments.  The form factors from different simulation models with
different cholesterol concentrations are compared to experimental data Fig.~\ref{FormFactors}. \\
\todo{Discussion to be finished when the data is solid.}

 \begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/FormFactors.eps}
  \caption{\label{FormFactors}
    Form factors from simulations and experiments.
  }

@@

Simulated systems are listed in Table \ref{systemsCHOL} and the full simulation
details are given in references or in supplementary material.

@@
